Declarative Integration of Interactive 3D Graphics into the World-Wide Web:\\Principles, Current Approaches, and Research Agenda

### 2013-04-29 Web3D2013-Dec3D/paper.tex
New image presenting debugging
Minor updates

It is also worth mentioning that the Web applications shown in Figures \ref{fig:chApps} and \ref{fig:cadApps} were not realized by graphics programmers but mainly by Web developers, using standard DOM scripting as well as JavaScript frameworks like jQuery \cite{jquery}. For example, to switch to a certain viewpoint when the user clicks onto the arrow button in the walkthrough application shown in Figure \ref{fig:chApps} (left), the application developer simply needs to update the 'bind' attribute of the "viewpoint" tag by calling its \emph{setAttribute} method.
Likewise, when selecting an element in the browser-based CAD and product structure viewer shown in the middle of Figure \ref{fig:cadApps}, the corresponding parts in the tree view are highlighted and vice versa.

@@

CSS 3D Transforms~\cite{webkit3DCSS} should be used, to define \textit{transformations}~(3) on the hierarchical structure as well as on other 3D elements. This example illustrates the three later concepts:

@@

In general, a tight \textit{CSS integration}~(4) is a major concept: existing CSS properties should be used where applicable (e.g., transform, opacity, color, etc.). In addition, a set of 3D-specific CSS properties (e.g. for defining the appearance) needs to be defined. However, custom CSS properties are not yet supported by any Web Browser, though finally this seems to change now \cite{CSSCustom}.

@@

Dec3D should support highly dynamic scenes. Using the DOM API to modify the scene data is not sufficient for many use cases, e.g.\ dynamic meshes, image processing, compression, etc. Thus we need a concept to support efficient data processing that allows to map computations to available hardware. These should be expressed as \textit{operators}~(10) on the data containers above.

@@

\textit{Lights}~(13) are shader parameters in OpenGL/ WebGL \cite{WebGL12}. In contrast, for Dec3D we aim a representation to provide a intuitive way to define lights (and thereby also shadows) and to not obstruct more advanced future concepts.
Likewise, \textit{viewpoints}~(14) can serve as link to certain points of interest within the document. However, it is still discussable if some pre-defined \textit{camera navigation}~(15) modes -- to get directly started without having to modify the camera matrix with a special script -- should be part of the essential profile.

@@

% http://www.w3.org/community/declarative3d/wiki/Level_of_Integration/
We also propose a measure for the Level of Integration~(LOI) for 3D graphics into the W3C technology stack (DOM, CSS, etc.) to explain existing and possible integration levels~(Figure \ref{fig:polyInt}). Level 0 is the classical integration using plug-ins. Many X3D implementations as well as Adobe's Flash and Unity 3D use the plug-in model to integrate their rendering engine into the browser. LOI 1 provides a dedicated element in the DOM in combination with an API. Currently  WebGL and the \verb|<canvas>| provide this level of integration. The next level (LOI 2) has the scene description not in JavaScript but integrated in the DOM. Levels 3 and 4 provide an even better integration with CSS and easy debugging of 3D scenes. This requires having a set of 3D-specific CSS properties or even a full shader description via CSS as well as a deeper integration into debugging tools like Chrome Developer Tools or Firebug.

@@

\paragraph{Encourage Participation}
All relevant stakeholders, e.g. developers, designers, researchers, 3D artists, industry professionals, representatives of standards organizations, accessibility experts, and user-agent implementers, are encouraged to participate in this group. Participants should be willing to actively develop and donate materials towards the group's deliverables, as well as attend the group's teleconferences and face-to-face meetings.

\paragraph{Clear Definition of Use Cases and Requirements}
The group needs to agree on a collection of relevant use cases, where embedding 3D data in HTML using a declarative approach provides significant benefit. Here, declarative approaches esp. seem relevant for the industry, since they tend to think in formats (such as STL, STEP/IGES or CATIA) not in APIs.
Each use case should explore how publishers and consumers benefit from Dec3D. From these use cases, the group needs to derive and prioritize the different required dimensions for the Dec3D technical specification \cite{JankowskiDec3D2012,LeFeuvreDec3D2012}.

\paragraph{Clear Technical Specification}
The next step is the creation of the clear, detailed and extensible technical specification of the implementation concepts and features necessary to cover a majority of useful requirements. Measurable properties need to be defined to quantitatively and/or qualitatively evaluate the achieved solution, document the pros and cons of each, and demonstrate that, based on the above analysis, there is a good chance of success in creating a W3C standard for Declarative 3D for the Web.

@@

\section*{Acknowledgements}
The research presented in this paper has been supported by the EU projects V-Must (\url{http://v-must.net}), VERVE (\url{http://www.verveconsortium.eu}) and FI-CONTENT, the BMBF project EMERGENT, the Intel Visual Computing Institute, the SFI's Lion2 and EI's Copernicus projects, and the Villes Transparentes project.

>>>>

It is also worth mentioning that the Web applications shown in Figures \ref{fig:chApps} and \ref{fig:cadApps} were not realized by graphics programmers, but mainly by Web developers using standard DOM scripting as well as JavaScript frameworks like jQuery \cite{jquery}. For example, to switch to a certain viewpoint when the user clicks onto the arrow button in the walkthrough application shown in Figure \ref{fig:chApps} (left), the application developer simply needs to update the 'bind' attribute of the "viewpoint" tag by calling its \emph{setAttribute} method.
Likewise, when selecting an element in the browser-based CAD and product structure viewer shown in the middle of Figure \ref{fig:cadApps}, the corresponding parts in the tree view are highlighted and vice versa.

@@

CSS 3D Transforms~\cite{webkit3DCSS} should be used to define \textit{transformations}~(3) on the hierarchical structure as well as on other 3D elements. This example illustrates the three following concepts:

@@

In general, a tight \textit{CSS integration}~(4) is a major concept: existing CSS properties should be used where applicable (e.g., transform, opacity, color, etc.). In addition, a set of 3D-specific CSS properties (e.g. for defining the appearance) needs to be defined. However, custom CSS properties are not yet supported by any Web browser, though finally this seems to change now \cite{CSSCustom}.

@@

Dec3D should support highly dynamic scenes. Using the DOM API to modify the scene data is not sufficient for many use cases, e.g.\ dynamic meshes, image processing, compression, etc. Thus, we need a concept to support efficient data processing that allows to map computations to available hardware. These should be expressed as \textit{operators}~(10) on the data containers above.

@@

\textit{Lights}~(13) are shader parameters in OpenGL/ WebGL \cite{WebGL12}. In contrast, for Dec3D we aim a representation to provide a intuitive way to define lights (and thereby also shadows) and to not obstruct more advanced future concepts.
Likewise, \textit{viewpoints}~(14) can serve as links to certain points of interest within the document. However, it is still discussable if some pre-defined \textit{camera navigation}~(15) modes -- to get directly started without having to modify the camera matrix with a special script -- should be part of the essential profile.

@@

% http://www.w3.org/community/declarative3d/wiki/Level_of_Integration/
We also propose a measure for the Level of Integration~(LOI) for 3D graphics into the W3C technology stack (DOM, CSS, etc.) to explain existing and possible integration levels~(Figure \ref{fig:polyInt}). Level 0 is the classical integration using plug-ins. Many X3D implementations as well as Adobe's Flash and Unity 3D use the plug-in model to integrate their rendering engine into the browser. LOI 1 provides a dedicated element in the DOM in combination with an API. Currently  WebGL and the \verb|<canvas>| provide this level of integration. The next level (LOI 2) has the scene description not in JavaScript but integrated in the DOM. Levels 3 and 4 provide an even better integration with CSS and easy debugging of 3D scenes. This requires having a set of 3D-specific CSS properties or even a full shader description via CSS as well as a deeper integration into debugging tools like Chrome Developer Tools or Firebug (for an example see Figure \ref{fig:debugger} and \cite{SonsSlusallekDec3D2012}).

\begin{figure}
  \centering
  \includegraphics[width=0.9\columnwidth]{images/museum.png}
  \caption{LOI 3 in the native XML3D implementation: Debugging of a 3D scene using the Chrome developer toolbar. Note also that materials are attached to meshes using the 'shader' CSS property.}
  \label{fig:debugger}
  %\vspace{-0.2cm}
\end{figure}

@@

\paragraph{Encourage Participation}
All relevant stakeholders, e.g. developers, designers, researchers, 3D artists, industry professionals, representatives of standards organizations, accessibility experts, and user-agent implementers, are encouraged to participate in this group. Participants should be willing to actively develop and donate materials towards the group's deliverables, as well as attend the group's teleconferences and face-to-face meetings.
\vspace{-0.1cm}

\paragraph{Clear Definition of Use Cases and Requirements}
The group needs to agree on a collection of relevant use cases, where embedding 3D data in HTML using a declarative approach provides significant benefit. Here, declarative approaches esp. seem relevant for the industry, since they tend to think in formats (such as STL, STEP/IGES or CATIA) not in APIs.
Each use case should explore how publishers and consumers benefit from Dec3D. From these use cases, the group needs to derive and prioritize the different required dimensions for the Dec3D technical specification \cite{JankowskiDec3D2012,LeFeuvreDec3D2012}.
\vspace{-0.1cm}

\paragraph{Clear Technical Specification}
The next step is the creation of the clear, detailed and extensible technical specification of the implementation concepts and features necessary to cover a majority of useful requirements. Measurable properties need to be defined to quantitatively and/or qualitatively evaluate the achieved solution, document the pros and cons of each, and demonstrate that, based on the above analysis, there is a good chance of success in creating a W3C standard for Declarative 3D for the Web.
\vspace{-0.1cm}

@@

\section*{Acknowledgements}
The research presented in this paper has been supported by the EU projects V-Must (\url{http://v-must.net}), VERVE (\url{http://www.verveconsortium.eu}) and FI-CONTENT, the BMBF project EMERGENT, the Intel Visual Computing Institute, the SFI's Lion2 and EI's Copernicus (\url{http://copernicus.deri.ie}) projects, and the Villes Transparentes project.
